---
title: "Analysis - Heme 1000"
output: html_notebook
---

# Data Preprocessing

```{r}

library(SpatialExperiment)
library(rtracklayer)
library(lobstr)
library(readr)
library(scater)
library(ggspavis)
library(sva)
library(plyr)

```

## Import data as SpatialExperiment object

```{r}

spe <- SpatialExperiment::read10xVisium(
  samples = "heme_1000",
  sample_id = "heme_1000",
  type = c("HDF5", "sparse"),
  data = "filtered",
  images = "lowres", 
  load = TRUE
)

```


Preso da SpatialLIBD (vedere bene!)

```{r}
# continuous variable containing total number of counts for each sample prior to filtering any genes
spe$sum_umi <- colSums(counts(spe))

# continuous variable containing the number of genes that have at least 1 count
spe$sum_gene <- colSums(counts(spe) > 0)
```


Reference genome alignment 

```{r}
# gene annotation - reference genome GRCm38.p6 http://ftp.ebi.ac.uk/pub/databases/gencode/Gencode_mouse/release_M23/gencode.vM23.annotation.gtf.gz

gtf <-                                           
  rtracklayer::import(
    "C:/Users/mason/OneDrive/Desktop/RGDIST/RGDIST 2.0/Control/gencode.vM23.annotation.gtf"
  )

gtf <- gtf[gtf$type == "gene"]                   
gtf$gene_id <- gsub("\\..*", "", gtf$gene_id)    
names(gtf) <- gtf$gene_id  

match_genes <- match(rownames(spe), gtf$gene_id) 
table(is.na(match_genes))
spe <- spe[!is.na(match_genes), ]                
match_genes <- match_genes[!is.na(match_genes)]
mcols(gtf) <- mcols(gtf)[, c("source", "type", "gene_id", "gene_name", "gene_type")]  
rowRanges(spe) <- gtf[match_genes]               

rowData(spe)$gene_search <- paste0(
  rowData(spe)$gene_name, "; ", rowData(spe)$gene_id
)

```


## Anatomy info and injection site coords 

```{r}

spots <- read_csv("heme_1000/outs/spatial/tissue_positions_list.csv", col_names = FALSE)
names(spots) <- c("Barcode", "in_tissue", "arrary_row", "array_col", "pixel_row", "pixel_col")
spots <- subset(spots, in_tissue == 1) # here we remove spots not covered by tissue


anatomy <- merge(read_csv("heme_1000/heme_1000_anatomy.csv"),
                 read_csv("heme_1000/heme_1000_injection_site.csv"),
                 by = "Barcode")

spots <- merge(spots, anatomy, by = "Barcode")


# Add info to spe object
spe$anatomy <- spots$anatomy
spe$inj_site <- spots$injection_site

```


## Distances from injection site

```{r}
# calcolo la distanza utilizzando le coordinate dello spot centrale del sito di iniezione 
# array_row = 48 ; array_col = 64
# spots[which(spots$arrary_row == 48 & spots$array_col == 64) , c(5,6)]
# pixel_row = 4828 ; pixel_col = 3620
spots$distance <- NA

for(i in 1:nrow(spots)){
  spots[i,"distance"] <- sqrt((spots[i,5] - 5118)^2 + (spots[i,6] - 3867)^2)
}

colData(spe)$inj_site_distance <- spots$distance

```


## Quality Control and data filtering

Mitochondrial genes

```{r}
# identify mitochondrial genes
is_mito <- grepl("(^MT-)|(^mt-)", rowData(spe)$gene_name)
table(is_mito)

# calculate per-spot QC metrics and store in colData
spe <- addPerCellQC(spe, subsets = list(mito = is_mito))

# histogram of numbers of expressed genes
hist(colData(spe)$detected, breaks = 20)

# Remove mitochondrial genes
spe <- spe[!is_mito, ]
```


NAs spots: spots that are not assigned to any anatomical cluster. This clustering was manually made by the dataset authors, some spots were not assigned to any cluster and since this information is crucial for our further analysis, we decided to remove them. 

```{r}
# we have some NAs in our anatomy info so we need to remove them
NA_spot <- c(which(is.na(spots[,"anatomy"])))  
spe <- spe[, -NA_spot, drop = FALSE]
spots <- spots[-NA_spot, ]

```


We also need to remove genes that are not expressed at all across the whole tissue. 

```{r}
# remove not expressed genes
no_expr <- which(rowSums(counts(spe)) == 0)           
length(no_expr) / nrow(spe) * 100  # percentage of not expressed genes                     
spe <- spe[-no_expr, , drop = FALSE]
```


```{r}
summary(spe$sum)

if (any(spe$sum == 0)) {
  spots_no_counts <- which(spe$sum == 0)
  ## Number of spots with no counts
  print(length(spots_no_counts))
  ## Percent of spots with no counts
  print(length(spots_no_counts) / ncol(spe) * 100)
  spe <- spe[, -spots_no_counts, drop = FALSE]
}
```

In order to compute correlations we need to have at least three counts for each gene. We consider as no relevant all those genes that have less than 3 counts.

```{r}
# Check the number of spots in which a gene is expressed and removal of genes expressed in less than 3 spots

h1000_counts <- as.matrix(assays(spe)$counts)

no_rel <- c()
n <- 0

for(i in 1:nrow(h1000_counts)){
  for(j in 1:ncol(h1000_counts)){
    if(h1000_counts[i,j] != 0){
      n = n + 1
    }
  }
  if(n <= 2){
    no_rel = c(no_rel, i)
  }
  n = 0
}

remove(h1000_counts)
spe <- spe[-no_rel, , drop = FALSE]

```


## Normalization of counts

In our object spe the only available assay is "counts" (assayNames(spe)). Normalization is an essential step in an RNA-Seq analysis, in which the read count matrix is transformed to allow for meaningful comparisons of counts across samples.

```{r}
spe <- scuttle::logNormCounts(spe)
```


## Batch Correction on anatomy cluster 

```{r}

h1000_counts <- as.matrix(assays(spe)$counts)

spots$batch <- as.factor(revalue(spots$anatomy,
                         c("caudate_putamen" = 1, 
                           "cortex" = 2, 
                           "thalamus" = 3, 
                           "globus_pallidus" = 4, 
                           "plexus" = 5, 
                           "hypothalamus" = 6, 
                           "corpus_callosum" = 7)))


BCcounts <- ComBat_seq(counts = h1000_counts,
                              batch = spots$batch,
                              group = NULL,
                              covar_mod = NULL,
                              full_mod = FALSE)

assays(spe)$BCcounts <- BCcounts

# Also the corrected counts must be normalized
spe <- scuttle::logNormCounts(x = spe,
                              assay.type = "BCcounts",
                              name = "BClogcounts")

```


## Dimensionality Reduction: UMAP 

```{r}
# No corrected data UMAP
spe <- runUMAP(spe, 
               exprs_values = "logcounts", 
               name = "UMAP_noBC")

plotReducedDim(object = spe, 
               dimred = "UMAP_noBC", 
               colour_by = "anatomy")


# Batch Corrected data UMAP

spe <- runUMAP(spe, 
               exprs_values = "BClogcounts", 
               name = "UMAP_BClog")

plotReducedDim(object = spe, 
               dimred = "UMAP_BClog", 
               colour_by = "anatomy")

```


## Correlations and p-values

```{r}
# BEFORE batch effect removal
h1000_logcounts <- as.matrix(assays(spe)$logcounts)
distances <- spe$inj_site_distance

cor_before <- c()
pval_before <- c()

for(i in 1:nrow(h1000_logcounts)) {
  zero <- which(h1000_logcounts[i,] == 0)
  if(!isEmpty(zero)){
    cor_before[i] <- cor(distances[-zero], h1000_logcounts[i,-zero], method = "spearman")
    test <- cor.test(distances[-zero], h1000_logcounts[i,-zero], method = "spearman")
    pval_before[i] <- test$p.value
  }else{
    if(isEmpty(zero)){
      cor_before[i] <- cor(distances, h1000_logcounts[i,], method = "spearman")
      test <- cor.test(distances, h1000_logcounts[i,], method = "spearman")
      pval_before[i] <- test$p.value
    }
  }
}

names(cor_before) <- rownames(rowData(spe))
names(pval_before) <- rownames(rowData(spe))

hist(cor_before)
hist(abs(cor_before))

# p-value correction
padjust_bf <- p.adjust(pval_before, method="fdr")

```


```{r}
# AFTER batch effect removal
BClog <- as.matrix(assays(spe)$BClogcounts)
cor_after <- c()
pval_after <- c()

for(i in 1:nrow(BClog)) {
  zero <- which(BClog[i,] == 0)
  if(!isEmpty(zero)){
    cor_after[i] <- cor(distances[-zero], BClog[i,-zero], method = "spearman")
    test <- cor.test(distances[-zero], BClog[i,-zero], method = "spearman")
    pval_after[i] <- test$p.value
  }else{
    if(isEmpty(zero)){
      cor_after[i] <- cor(distances, BClog[i,], method = "spearman")
      test <- cor.test(distances, BClog[i,], method = "spearman")
      pval_after[i] <- test$p.value
    }
  }
}

names(cor_after) <- rownames(rowData(spe))
names(pval_after) <- rownames(rowData(spe))

hist(cor_after,
     main = paste("Histogram of" , "heme 1000 correlations"),
     col = "lightblue",
     xlab = "Correlation")
hist(abs(cor_after))

# p-value correction
padjust_af <- p.adjust(pval_after, method="fdr")

```

```{r}
# Add information to spe
rowData(spe)$cor_before <- cor_before
rowData(spe)$pvalue_before <- pval_before
rowData(spe)$pvalue_fdr_before <- padjust_bf

rowData(spe)$cor_after <- cor_after
rowData(spe)$pvalue_after <- pval_after
rowData(spe)$pvalue_fdr_after <- padjust_af
```

## Ranking

```{r}

rowData(spe)$rank_before <- rank(padjust_bf, ties.method = "min")
rowData(spe)$rank_after <- rank(padjust_af,  ties.method = "min")

```


# Extract significant genes

```{r}
cor_genes_bf <- which(cor_before >= 0.5 & padjust_bf < 0.01)
anticor_genes_bf <- which(cor_before <= -0.5 & padjust_bf < 0.01)

col <- rep("gray", nrow(rowData(spe)))
col[cor_genes_bf] <- "pink"
col[anticor_genes_bf] <- "green"

gene_name <- rowData(spe)$gene_name

plot(x = cor_before,
     y = -log(padjust_bf),
     xlab = "Correlation",
     ylab = "-log p.values",
     type = "p",
     col = col,
     pch = 20,
     main = "Before batch correction"
)
abline(v=-0.5, col="black", lty=2)
abline(v=0.5, col="black", lty=2)
abline(h=-log(0.01), col="black", lty=2)
x = c(cor_before[anticor_genes_bf], cor_before[cor_genes_bf])
y = c(padjust_bf[anticor_genes_bf], padjust_bf[cor_genes_bf])
text(x, -log(y), 
     labels = c(gene_name[anticor_genes_bf], gene_name[cor_genes_bf]),
     cex = 0.4)


# After BC
cor_genes_af <- which(cor_after >= 0.35 & padjust_af <= 0.01)
anticor_genes_af <- which(cor_after <= -0.35 & padjust_af <= 0.01)

col <- rep("gray", nrow(rowData(spe)))
col[cor_genes_af] <- "pink"
col[anticor_genes_af] <- "green"

plot(x = cor_after,
     y = -log(padjust_af),
     xlab = "Correlation",
     ylab = "-log p.values",
     type = "p",
     col = col2,
     pch = 20,
     main = "After batch correction"
)
abline(v=-0.35, col="black", lty=2)
abline(v=0.35, col="black", lty=2)
abline(h=-log(0.01), col="black", lty=2)
x = c(cor_after[anticor_genes_af], cor_after[cor_genes_af])
y = c(padjust_af[anticor_genes_af], padjust_af[cor_genes_af])
text(x, -log(y), 
     labels = c(gene_name[anticor_genes_af], gene_name[cor_genes_af]),
     cex = 0.4)

```
