---
title: "Analysis - Control Sample"
output: html_notebook
---

## Data preprocessing
### Import data as SpatialExperiment object
```{r}
library(SpatialExperiment)
library(rtracklayer)
library(lobstr)

spe <- SpatialExperiment::read10xVisium(
  samples = "control",
  sample_id = "control",
  type = c("HDF5", "sparse"),
  data = "filtered",
  images = "lowres", 
  load = TRUE
)

# continuous variable containing total number of counts for each sample prior to filtering any genes
spe$sum_umi <- colSums(counts(spe))
# continuous variable containing the number of genes that have at least 1 count
spe$sum_gene <- colSums(counts(spe) > 0)

# gene annotation - reference genome GRCm38.p6 http://ftp.ebi.ac.uk/pub/databases/gencode/Gencode_mouse/release_M23/
gtf <-                                           
  rtracklayer::import(
    "gencode.vM23.annotation.gtf"
  )

gtf <- gtf[gtf$type == "gene"]                   
gtf$gene_id <- gsub("\\..*", "", gtf$gene_id)    
names(gtf) <- gtf$gene_id                        

match_genes <- match(rownames(spe), gtf$gene_id) 
table(is.na(match_genes))
spe <- spe[!is.na(match_genes), ]                
match_genes <- match_genes[!is.na(match_genes)]
mcols(gtf) <- mcols(gtf)[, c("source", "type", "gene_id", "gene_name", "gene_type")]  
rowRanges(spe) <- gtf[match_genes]               

rowData(spe)$gene_search <- paste0(
  rowData(spe)$gene_name, "; ", rowData(spe)$gene_id
)
```

### Anatomy info and injection site coords
```{r}
library(readr)

control_anatomy <- read_csv("control/control_anatomy.csv")
control_injection_site <- read_csv("control/control_injection_site.csv")

control_ana_inj <- merge(control_anatomy, control_injection_site, by = "Barcode")
remove(control_anatomy, control_injection_site)

spe$anatomy <- control_ana_inj$anatomy
spe$inj_site <- control_ana_inj$injection_site
```

### Distances from injection site
```{r}
tissue_positions_list_control <- read_csv("control/outs/spatial/tissue_positions_list.csv", col_names = FALSE)
names(tissue_positions_list_control) <- c("Barcode", "in_tissue", "arrary_row", "array_col", "pixel_row", "pixel_col")
tissue_positions_list_control <- subset(tissue_positions_list_control, in_tissue == 1)
tissue_positions_list_control <- tissue_positions_list_control[,-c(5,6)]
control_positions <- merge(tissue_positions_list_control, control_ana_inj, by = "Barcode")
remove(tissue_positions_list_control)

# calcolo la distanza utilizzando le coordinate dello spot centrale del sito di iniezione 
# x = 44 ; y = 70
control_positions$distance <- NA
for(i in 1:nrow(control_positions)){
  control_positions[i,7] <- sqrt((control_positions[i,3] - 44)^2 + (control_positions[i,4] - 70)^2)
}

colData(spe)$inj_site_distance <- control_positions$distance
```

### Quality Control and data filtering ----
```{r}
library(scater)
library(ggspavis)

# identify mitochondrial genes
is_mito <- grepl("(^MT-)|(^mt-)", rowData(spe)$gene_name)
table(is_mito)

# calculate per-spot QC metrics and store in colData
spe <- addPerCellQC(spe, subsets = list(mito = is_mito))

# histogram of numbers of expressed genes
hist(colData(spe)$detected, breaks = 20)

# Remove mitochondrial genes
spe <- spe[!is_mito, ]

# we have some NAs in our anatomy info so we need to remove them
NA_spot <- c(which(is.na(control_positions[,5]))) 
spe <- spe[,-NA_spot, drop = FALSE]

# remove not expressed genes
no_expr <- which(rowSums(counts(spe)) == 0)           
length(no_expr) / nrow(spe) * 100                     
spe <- spe[-no_expr, , drop = FALSE]

summary(spe$sum)

if (any(spe$sum == 0)) {
  spots_no_counts <- which(spe$sum == 0)
  ## Number of spots with no counts
  print(length(spots_no_counts))
  ## Percent of spots with no counts
  print(length(spots_no_counts) / ncol(spe) * 100)
  spe <- spe[, -spots_no_counts, drop = FALSE]
}

# Compute the number of 0s values
control_counts <- assays(spe)$counts
control_counts <- as.matrix(control_counts)

zero_x_gene <- c()
for(i in 1:nrow(control_counts)){
  zero_x_gene[i] = sum(control_counts[i,] == 0)
}

no_zero_x_gene <- 2209 - zero_x_gene 
rowData(spe)$zero_x_gene <- zero_x_gene
rowData(spe)$no_zero_x_gene <- no_zero_x_gene

# Check the number of spots in which a gene is expressed and removal of genes expressed in less than 3 spots

no_rel <- c()
n <- 0

for(i in 1:nrow(control_counts)){
  for(j in 1:ncol(control_counts)){
    if(control_counts[i,j] != 0){
      n = n + 1
    }
  }
  if(n <= 2){
    no_rel = c(no_rel, i)
  }
  n = 0
}

remove(control_counts)
spe <- spe[-no_rel, , drop = FALSE]
control_positions <- control_positions[-NA_spot,]

# Normalization
spe <- scuttle::logNormCounts(spe)

remove(i,j,match_genes,is_mito,control_ana_inj,gtf,no_expr, no_rel,n, zero_x_gene, no_zero_x_gene)
```

### Batch Correction on anatomy cluster
```{r}
library(sva)

control_counts <- assays(spe)$counts
control_counts <- as.matrix(control_counts)

batch <- c()
for(i in 1:nrow(control_positions)){
  if(is.na(control_positions[i,5])){
    batch[i] <- NA
  }else{
    if(control_positions[i,5] == "caudate_putamen"){
      batch[i] <- 1
    }else{
      if(control_positions[i,5] == "cortex"){
        batch[i] <- 2
      }else{
        if(control_positions[i,5] == "thalamus"){
          batch[i] <- 3
        }else{
          if(control_positions[i,5] == "globus_pallidus"){
            batch[i] <- 4
          }else{
            if(control_positions[i,5] == "plexus"){
              batch[i] <- 5
            }else{
              if(control_positions[i,5] == "hypothalamus"){
                batch[i] <- 6
              }else{
                if(control_positions[i,5] == "corpus_callosum"){
                  batch[i] <- 7
                }
              }
            }
          }
        }
      }
    }
  }
}
control_positions$batch <- batch

BCcounts <- ComBat_seq(counts = control_counts,
                            batch = batch,
                            group = NULL,
                            covar_mod = NULL,
                            full_mod = FALSE)

assays(spe)$BC <- BCcounts

spe <- logNormCounts(x = spe,
                           assay.type = "BC",
                           name = "BClogcounts")
```

### Reduced dimension: UMAP
```{r}
library(scater)

# No corrected data UMAP
spe <- runUMAP(spe, exprs_values = "logcounts", name = "UMAP_noBC")
plotReducedDim(object = spe, dimred = "UMAP_noBC", colour_by = "anatomy")

# Batch Corrected data UMAP
spe <- runUMAP(spe, exprs_values = "BC", name = "UMAP_BC")
plotReducedDim(object = spe, dimred = "UMAP_BC", colour_by = "anatomy")
```

### Correlation and p.values
```{r}
# BEFORE batch effect removal
counts_log <- assays(spe)$logcounts
counts_log <- as.matrix(counts_log)
distance_c <- spe$inj_site_distance

# Calcolo la correlazione tra la distanza e l'espressione per ogni gene e il relativo p.value
cor_b <- c()
pval_b <- c()

for(i in 1:nrow(counts_log)) {
  zero <- which(counts_log[i,] == 0)
  if(!isEmpty(zero)){
    cor_b[i] <- cor(distance_c[-zero], counts_log[i,-zero], method = "spearman")
    test <- cor.test(distance_c[-zero], counts_log[i,-zero], method = "spearman")
    pval_b[i] <- test$p.value
  }else{
    if(isEmpty(zero)){
      cor_b[i] <- cor(distance_c, counts_log[i,], method = "spearman")
      test <- cor.test(distance_c, counts_log[i,], method = "spearman")
      pval_b[i] <- test$p.value
    }
  }
}

names(cor_b) <- rownames(rowData(spe))
names(pval_b) <- rownames(rowData(spe))

hist(cor_b)
hist(abs(cor_b))

# Correzione pvalue
padjust_bf <- p.adjust(pval_b, method="fdr")

# AFTER batch effect removal
BC_log <- assays(spe)$BClogcounts
BC_log <- as.matrix(BC_log)
cor_a <- c()
pval_a <- c()

for(i in 1:nrow(BC_log)) {
  zero <- which(BC_log[i,] == 0)
  if(!isEmpty(zero)){
    cor_a[i] <- cor(distance_c[-zero], BC_log[i,-zero], method = "spearman")
    test <- cor.test(distance_c[-zero], BC_log[i,-zero], method = "spearman")
    pval_a[i] <- test$p.value
  }else{
    if(isEmpty(zero)){
      cor_a[i] <- cor(distance_c, BC_log[i,], method = "spearman")
      test <- cor.test(distance_c, BC_log[i,], method = "spearman")
      pval_a[i] <- test$p.value
    }
  }
}

names(cor_a) <- rownames(rowData(spe))
names(pval_a) <- rownames(rowData(spe))

hist(cor_a)
hist(abs(cor_a))

# correzione pvalue
padjust_af <- p.adjust(pval_a, method="fdr")

# Add informations to spe
rowData(spe)$cor_before <- cor_b
rowData(spe)$pvalue_before <- padjust_bf
rowData(spe)$pvalue_before_NOadj <- pval_b

rowData(spe)$cor_after <- cor_a
rowData(spe)$pvalue_after <- padjust_af
rowData(spe)$pvalue_after_NOadj <- pval_a
```
```{r}
spots <- control_positions
remove(control_positions)

genes <- rowData(spe)
genes <- as.data.frame(genes)
genes <- genes[,-c(1,2)]
```

```{r}
remove(BC_log, BCcounts, control_counts, test, batch, cor_a, cor_b, pval_a, pval_b, padjust_af, padjust_bf, zero, i, distance_c, counts_log)
```



## Significant genes: correlated and anticorrelated

Extract significant genes

```{r}
# Before BC
pval_bf <- rowData(spe)$pvalue_before
cor_bf <- rowData(spe)$cor_before

upg_bf <- which(cor_bf >= 0.5 & pval_bf < 0.01)
dng_bf <- which(cor_bf <= -0.5 & pval_bf < 0.01)

col <- rep("gray", nrow(genes))
col[upg_bf] <- "lightsalmon"
col[dng_bf] <- "seagreen1"

gene_name <- rowData(spe)$gene_name
nom <- c(gene_name[dng_bf], gene_name[upg_bf])

plot(x = cor_bf,
     y = -log(pval_bf),
     xlab = "Correlation",
     ylab = "-log p.values",
     type = "p",
     col = col,
     pch = 20,
     main = "Before batch correction"
)
abline(v=-0.5, col="black", lty=2)
abline(v=0.5, col="black", lty=2)
abline(h=-log(0.01), col="black", lty=2)
x = c(cor_bf[dng_bf], cor_bf[upg_bf])
y = c(pval_bf[dng_bf], pval_bf[upg_bf])
text(x, -log(y), 
     labels = nom,
     cex = 0.4)

```

```{r}
# After BC
pval_af <- rowData(spe)$pvalue_after
cor_af <- rowData(spe)$cor_after

upg_af <- which(cor_af >= 0.35 & pval_af <= 0.01)
dng_af <- which(cor_af <= -0.35 & pval_af <= 0.01)

col2 <- rep("gray", nrow(genes))
col2[upg_af] <- "lightsalmon"
col2[dng_af] <- "seagreen1"

# vettore nomi
nom2 <- c(gene_name[dng_af], gene_name[upg_af])

plot(x = cor_af,
     y = -log(pval_af),
     xlab = "Correlation",
     ylab = "-log p.values",
     type = "p",
     col = col2,
     pch = 20,
     main = "After batch correction"
)
abline(v=-0.35, col="black", lty=2)
abline(v=0.35, col="black", lty=2)
abline(h=-log(0.01), col="black", lty=2)
x = c(cor_af[dng_af], cor_af[upg_af])
y = c(pval_af[dng_af], pval_af[upg_af])
text(x, -log(y), 
     labels = nom2,
     cex = 0.4)
```

```{r}
## nostri sig genes
cor_before <- genes[upg_bf,]
anticor_before <- genes[dng_bf,]

cor_after <- genes[upg_af,]
anticor_after <- genes[dng_af,]
```

```{r}
remove(col, col2, cor_af, cor_bf, dng_af, dng_bf, upg_af, upg_bf, gene_name, nom, nom2, x, y, pval_af, pval_bf)
```

```{r}

write(anticor_before$gene_id, "anticor_bf_c.txt")
write(anticor_after$gene_id, "anticor_af_c.txt")
write(cor_before$gene_id, "cor_bf_c.txt")
write(cor_after$gene_id, "cor_af_c.txt")

```


\newpage
## Ground truth genes

```{r}
library(readr)
heme_response_downgenes <- read_csv("heme_response_downgenes.csv")
heme_response_upgenes <- read_csv("heme_response_upgenes.csv")
```
```{r}
# dove si trovano i loro geni nei nostri dati ----
## down genes ---

dngenes_names <- heme_response_downgenes$gene_name
heme_dn_position <- c()
for(i in 1:6){
  heme_dn_position[i] = which(genes[,2] == dngenes_names[i])
}


## up genes ---

## TGM1 GENE IS NOT IN OUR DATA, IT HAS BEEN REMOVED DURING DATA PREPROCESSING. Same for Ptx3 e Arg1

upgenes_names <- heme_response_upgenes$gene_name
heme_up_position <- c()
for(i in 1:59){
  heme_up_position[i] = which(genes[,2] == upgenes_names[i])
}


```



\newpage

## Ranking

Ordiniamo i risultati in base al p-value della correlazione di Spearman, dal più piccolo al più grande

```{r}

genes$rank_before <- rank(genes$pvalue_before, ties.method = "min")
genes$rank_after <- rank(genes$pvalue_after, ties.method = "min")

```


## Baseline

```{r}
# baseline method
# zg <- rowData(spe)$zero_x_gene
# rowData(spe)$zero_x_gene <- zg
# no_zero_x_gene <- 2209 - zg
# rowData(spe)$no_zero_x_gene <- no_zero_x_gene

no_zero_x_gene <- rowData(spe)$no_zero_x_gene

# before
logcounts_matrix <- assays(spe)$logcounts
logcounts_matrix <- as.matrix(logcounts_matrix)
baseline_before <- rowSums(logcounts_matrix)/no_zero_x_gene
mean_expr_before <- mean(baseline_before)

dist_mean_before <- c()
for(i in 1:length(baseline_before)){
  dist_mean_before[i] = abs(baseline_before[i] - mean_expr_before)
}

rank_baseline_before <- rank(dist_mean_before, ties.method = "min")
rank_baseline_before <- 16146 - rank_baseline_before

rowData(spe)$baseline_before <- baseline_before
rowData(spe)$rank_baseline_before <- rank_baseline_before

# after
BClogcounts <- assays(spe)$BClogcounts
BClogcounts <- as.matrix(BClogcounts)
baseline_after <- rowSums(BClogcounts)/no_zero_x_gene
mean_expr_after <- mean(baseline_after)

dist_mean_after <- c()
for(i in 1:length(baseline_after)){
  dist_mean_after[i] = abs(baseline_after[i] - mean_expr_after)
}

rank_baseline_after <- rank(dist_mean_after, ties.method = "min")
rank_baseline_after <- 16146 - rank_baseline_after

rowData(spe)$baseline_after <- baseline_after
rowData(spe)$rank_baseline_after <- rank_baseline_after

baseline <- genes[,-c(3,4,5,6,9,12,13,14)]

baseline$baseline_before <- baseline_before
baseline$rank_baseline_before <- rank_baseline_before
baseline$baseline_after <- baseline_after
baseline$rank_baseline_after <- rank_baseline_after

```


\newpage
## SVG
```{r}
library(spatialDE)

# preparazione input
x_cord <- spe$array_row
y_cord <- spe$array_col
spots <- colnames(spe)
sample_info <- matrix(data = NA, nrow = 2209, ncol = 2)
colnames(sample_info) <- c("x", "y")
rownames(sample_info) <- spots
sample_info[,1] <- x_cord
sample_info[,2] <- y_cord
sample_info <- as.data.frame(sample_info)
gene_id <- rowData(spe)$gene_id
expr_counts <- assays(spe)$counts
expr_counts <- as.matrix(expr_counts)
total_counts <- colSums(expr_counts)
sample_info$total_counts <- total_counts
coordinates <- sample_info[,-3]

# pipeline svg
expr_norm <- stabilize(expr_counts)
expr_resid <- regress_out(expr_norm, sample_info = sample_info)
results <- spatialDE::run(expr_resid, coordinates = coordinates)

head(results[order(results$qval), ])

de_results <- results[results$qval < 0.01, ] 

SVG_before <- results[,c("g", "LLR", "pval")]
rownames(SVG_before) <- c(1:16145)

rank_llr <- rank(SVG_before$LLR, ties.method = "min")
SVG_before$rank_llr_before <- 16146 - rank_llr
rownames(SVG_before) <- c(1:16145)

```

```{r}
# AFTER BATCH CORRECTION
sample_info_BC <- sample_info[,-3]
BCcounts <- assays(spe)$BC
BCcounts <- as.matrix(BCcounts)
total_counts_BC <- colSums(BCcounts)
sample_info_BC$total_counts <- total_counts_BC

expr_norm_BC <- stabilize(BCcounts)
expr_resid_BC <- regress_out(expr_norm_BC, sample_info = sample_info_BC)
results_BC <- spatialDE::run(expr_resid_BC, coordinates = coordinates)

# de_results_BC <- results_BC[results_BC$qval < 0.01, ] 

# head(results[order(results_BC$qval), ])

SVG_BC <- results_BC[,c("g", "LLR", "pval")]
rownames(SVG_BC) <- c(1:16145)

rank_llr_BC <- rank(SVG_BC$LLR, ties.method = "min")
SVG_BC$rank_llr_after <- 16146 - rank_llr_BC
rownames(SVG_BC) <- c(1:16145)

SVG <- merge(SVG_before, SVG_BC, by = "g")

# Posizioni dei geni ground truth
SVG_up <- c()
up_id <- heme_response_upgenes$gene_id
for (i in 1:59) {
  SVG_up[i] <- which(SVG$g == up_id[i])
}

SVG_dn <- c()
dn_id <- heme_response_downgenes$gene_id
for (i in 1:6) {
  SVG_dn[i] <- which(SVG$g == dn_id[i])
}
```


Il pvalue è calcolato sulla statistica LLR, per cui facciamo i rank di questi valori
```{r}
cor_pLLR <- cor(SVG_before$LLR, SVG_before$pval)

```



\newpage
## Methods comparison

```{r}
# Posizioni dei geni ground truth
# baseline_up <- c()
# up_id <- heme_response_upgenes$gene_id
# for (i in 1:59) {
#   baseline_up[i] <- which(baseline$gene_id == up_id[i])
# }
# 
# baseline_dn <- c()
# dn_id <- heme_response_downgenes$gene_id
# for (i in 1:6) {
#   baseline_dn[i] <- which(baseline$gene_id == dn_id[i])
# }
```


```{r}
par(mar=c(5, 4, 4, 8), xpd=TRUE)
boxplot(baseline$rank_baseline_before[c(heme_up_position, heme_dn_position)], 
        baseline$rank_baseline_after[c(heme_up_position, heme_dn_position)],
        SVG$rank_llr_before[c(SVG_up, SVG_dn)], SVG$rank_llr_after[c(SVG_up, SVG_dn)],
        genes$rank_before[c(heme_up_position, heme_dn_position)], genes$rank_after[c(heme_up_position, heme_dn_position)],
        names = c("no ComBat", "ComBat","no ComBat", "ComBat", "no ComBat", "ComBat"),
        xlab = "Method",
        ylab = "Rank",
        cex.axis = 0.8,
        cex.lab = 0.9,
        outline = F,
        # ylim = c(16150,1),
        main = "Rank of ground truth genes - methods compared",
        col = c("lightgoldenrod1", "lightgoldenrod1", "plum1", "plum1", "turquoise", "turquoise"))
legend("topleft", inset=c(1.01,0), legend = c("Baseline", "SVG", "Spearman"),
       fill = c("lightgoldenrod1", "plum1", "turquoise"), cex = 0.7,
       title = "Methods legend", title.cex = 0.7, bty = "n")
x <- rep(1:6, each = 65)
cols <- rep(c("gold2","gold2", "plum", "plum", "turquoise4", "turquoise4"), each = 65)
y <- c(baseline$rank_baseline_before[c(heme_up_position, heme_dn_position)], baseline$rank_baseline_after[c(heme_up_position, heme_dn_position)], 
       SVG$rank_llr_before[c(SVG_up, SVG_dn)], SVG$rank_llr_after[c(SVG_up, SVG_dn)],
       genes$rank_before[c(heme_up_position, heme_dn_position)], genes$rank_after[c(heme_up_position, heme_dn_position)])
points(jitter(x),y, col = cols)


par(mar=c(5, 4, 4, 8), xpd=TRUE)
boxplot(SVG$rank_llr_before[c(SVG_up, SVG_dn)], SVG$rank_llr_after[c(SVG_up, SVG_dn)],
        genes$rank_before[c(heme_dn_position, heme_up_position)], genes$rank_after[c(heme_dn_position, heme_up_position)],
        names = c("no ComBat", "ComBat", "no ComBat", "ComBat"),
        xlab = "Method",
        ylab = "Rank",
        cex.axis = 0.8,
        cex.lab = 0.9,
        outline = F,
        # ylim = c(15200,1),
        main = "Rank of ground truth genes - methods compared",
        col = c("plum1", "plum1", "turquoise3", "turquoise3"))
legend("topleft", inset=c(1.01,0), legend = c("SVG", "Spearman"),
       fill = c("plum1", "turquoise3"), cex = 0.7,
       title = "Methods legend", title.cex = 0.7, bty = "n")
x <- rep(1:4, each = 65)
cols <- rep(c("plum", "plum", "turquoise4", "turquoise4"), each = 65)
y <- c(SVG$rank_llr_before[c(SVG_up, SVG_dn)], SVG$rank_llr_after[c(SVG_up, SVG_dn)],
       genes$rank_before[c(heme_up_position, heme_dn_position)], genes$rank_after[c(heme_up_position, heme_dn_position)])
points(jitter(x),y, col = cols)
```


controllare pvalue 0.623713511 (rank: 8207)
plot : spearman trattato vs controllo

Rifare calcolo: spot min = 5
rimuovere geni espressi almeno 3-5

```{r}
rank.8207 <- which(genes[,"rank_after"] == 8207)
head(genes[rank.8207, c("cor_before", "pvalue_before", "cor_after", "pvalue_after")])

```

```{r}
# plot spearman: trattato vs controllo


heme_up_position <- c()
up_id <- heme_upgenes$id
for (i in 1:62) {
  heme_up_position[i] <- which(genes$gene_id == up_id[i])
}

heme_dn_position <- c()
dn_id <- heme_dngenes$id
for (i in 1:6) {
  heme_dn_position[i] <- which(genes$gene_id == dn_id[i])
}



par(mar=c(5, 4, 4, 8), xpd=TRUE)
boxplot(genes$rank_before[c(heme_dn_position, heme_up_position)], genes$rank_after[c(heme_dn_position, heme_up_position)],
        genes.control$rank_before[c(heme_dn_position_CONTROL, heme_up_position_CONTROL)], genes.control$rank_after[c(heme_dn_position_CONTROL, heme_up_position_CONTROL)],
        names = c("HEME noCB", "HEME CB", " CTRL noCB", "CTRL CB"),
        xlab = "",
        ylab = "Rank",
        cex.axis = 0.8,
        cex.lab = 0.9,
        outline = F,
        # ylim = c(15200,1),
        main = "Rank of ground truth genes - TREATED vs CTRL",
        col = c("darkolivegreen2", "darkolivegreen2", "hotpink1", "hotpink1"))
legend("topleft", inset=c(1.01,0), legend = c("heme", "ctrl"),
       fill = c("darkolivegreen2", "hotpink1"), cex = 0.7,
       title = "", title.cex = 0.7, bty = "n")


x2 <- rep(1:2, each = 68)
cols2 <- rep(c("darkolivegreen4", "darkolivegreen4"), each = 68)
y2 <- c(genes$rank_before[c(heme_up_position, heme_dn_position)], genes$rank_after[c(heme_up_position, heme_dn_position)])
points(jitter(x2),y2, col = cols2)

x1 <- rep(3:4, each = 65)
cols1 <- rep(c("hotpink4", "hotpink4"), each = 65)
y1 <- c(genes.control$rank_before[c(heme_dn_position_CONTROL, heme_up_position_CONTROL)], genes.control$rank_after[c(heme_dn_position_CONTROL, heme_up_position_CONTROL)])
points(jitter(x1),y1, col = cols1)
```

